### 2018-02-19 Physics Notes/Fields and their consequences/Magnetic Fields/Magnetic Fields.tex
Notes Update

\titlespacing\section{0pt}{14pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
\newlength\tindent
\setlength{\tindent}{\parindent}
\setlength{\parindent}{0pt}
\renewcommand{\indent}{\hspace*{\tindent}}

\pagestyle{fancy}
\fancyhf{}
\rhead{Sam Robbins 13SE}
\lhead{A Level Physics - Fields}
\rfoot{Page \thepage}

\begin{document}
\begin{center}
\underline{\huge Magnetic Fields}
\end{center}
\section{Charged Particles in magnetic fields}
Force is always perpendicular to the direction - no change in speed to the electrons.\\
\\
If electrons are confined to a wire $F=BIl$\\
\\
The same equation applies for many charges where:\\
\\
$I=\dfrac{Q}{t} \quad l=vt$\\
\\
$F=B\times\frac{Q}{t}\times vt$\\
\\
$F=BQv$\\
\\
\subsection{The hall probe}
This is used to measure the strength of magnetic fields. It consists of a thin slice of semiconductor material with a constant current flowing through it.\\
\\
The negative charges go to the bottom of the material and so the top is positively charged, this creates a potential difference.\\
\\
$F_B=F_E$\\
\\
$BQV=\dfrac{VQ}{d}$\\
\\
$V=Bvd$\\
\\
$V\propto B$\\
\\
By measuring the voltage across the probe the magnetic field strength can be measured.
\section{Charged particles in circular orbits}
\textbf{Force} will always be perpendicular to the direction of motion\\
No work is done by the magnetic field on the particles\\
\\
Direction changes but not speed\\
Constant kinetic energy\\
\\
Path is circular\\
Causes a centripetal acceleration\\
\\
\\
$$F=BQv=\dfrac{mv^2}{r}$$
$$r=\dfrac{mv}{BQ}$$
\newpage
\section{Applications}
\subsection{Cyclotron}
This is used in hospitals to produce high energy beams for radiotherapy.\\
\includegraphics[width=10cm]{cyclotron.jpg}\\
\begin{itemize}
\item Proton source in centre
\item Electric field in between C Cores to accelerate protons
\item Magnetic field causes curved path
\item r increases as v increases
\end{itemize}
\subsection{Mass Spectrometer}
This is used to analyse the type of atoms present in a sample\\
\\
\\
\begin{tikzpicture}
\draw[ultra thick, ->,postaction={decorate, decoration={text along path, raise=4pt, text align={align=center}, text={Light ion}, reverse path}}] (6.5,0) arc (0:180:2);
\draw[ultra thick, ->,postaction={decorate, decoration={text along path, raise=4pt, text align={align=center}, text={Heavy ion}, reverse path}}] (6.5,0) arc (0:180:3);
\draw[red,thick,fill=red] (0,0)rectangle (3,-0.25) node[below,midway] {Detector};
\draw[black,fill=black,thick] (5,0)rectangle (6,-0.25);
\draw[black,fill=black,thick] (7,0)rectangle (8,-0.25);
\draw[black,ultra thick] (6.5,0) -- (6.5,-2);
\end{tikzpicture}
\\
\begin{itemize}
\item Magnetic field out of diagram
\item Atoms in the sample are ionised
\item Directed in a narrow beam into the magnetic field
\item Radius of path depends on specific charge
\end{itemize}
$r=\dfrac{mv}{BQ}$\\
$r\propto\dfrac{1}{\frac{q}{m}}$
\subsubsection{Velocity selector}
\includegraphics[width=10cm]{velocity_selector.jpg}\\
Negative ions fall to the bottom surface, causing the top surface to have a positive charge, creating a potential difference. For an ion to continue on a straight path the force from the magnetic field up must be equal to the force from the electric field down.\\
\\
$F=BQv=EQ$\\
\\
$BQv=\dfrac{V}{d}Q$\\
\\
$v=\dfrac{V}{Bd}$\\
\\
This means that only particles of speed $v$ will travel through the gap

>>>>

\titlespacing\section{0pt}{14pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
\newlength\tindent
\setlength{\tindent}{\parindent}
\setlength{\parindent}{0pt}
\renewcommand{\indent}{\hspace*{\tindent}}
\hyphenpenalty=10000
\pagestyle{fancy}
\fancyhf{}
\rhead{Sam Robbins 13SE}
\lhead{A Level Physics - Fields}
\rfoot{Page \thepage}

\begin{document}
\begin{center}
\underline{\huge Magnetic Fields}
\end{center}
\section{Charged Particles in magnetic fields}
Force is always perpendicular to the direction - no change in speed to the electrons.\\
\\
If electrons are confined to a wire $F=BIl$\\
\\
The same equation applies for many charges where:
$$I=\frac{Q}{t} \quad l=vt$$
$$F=B\times\frac{Q}{t}\times vt$$
$$F=BQv$$
\subsection{The hall probe}
This is used to measure the strength of magnetic fields. It consists of a thin slice of semiconductor material with a constant current flowing through it.\\
\\
The negative charges go to the bottom of the material and so the top is positively charged, this creates a potential difference.
$$F_B=F_E$$
$$BQV=\frac{VQ}{d}$$
$$V=Bvd$$
$$V\propto B$$
By measuring the voltage across the probe the magnetic field strength can be measured.
\section{Charged particles in circular orbits}
\textbf{Force} will always be perpendicular to the direction of motion\\
No work is done by the magnetic field on the particles\\
\\
Direction changes but not speed\\
Constant kinetic energy\\
\\
Path is circular\\
Causes a centripetal acceleration\\
$$F=BQv=\dfrac{mv^2}{r}$$
$$r=\dfrac{mv}{BQ}$$
\newpage
\section{Applications}
\subsection{Cyclotron}
This is used in hospitals to produce high energy beams for radiotherapy.\\
\includegraphics[width=10cm]{cyclotron.jpg}\\
\begin{itemize}
\item Proton source in centre
\item Electric field in between C Cores to accelerate protons
\item Magnetic field causes curved path
\item r increases as v increases
\end{itemize}
\subsection{Mass Spectrometer}
This is used to analyse the type of atoms present in a sample\\
\\
\\
\begin{tikzpicture}
\draw[ultra thick, ->,postaction={decorate, decoration={text along path, raise=4pt, text align={align=center}, text={Light ion}, reverse path}}] (6.5,0) arc (0:180:2);
\draw[ultra thick, ->,postaction={decorate, decoration={text along path, raise=4pt, text align={align=center}, text={Heavy ion}, reverse path}}] (6.5,0) arc (0:180:3);
\draw[red,thick,fill=red] (0,0)rectangle (3,-0.25) node[below,midway] {Detector};
\draw[black,fill=black,thick] (5,0)rectangle (6,-0.25);
\draw[black,fill=black,thick] (7,0)rectangle (8,-0.25);
\draw[black,ultra thick] (6.5,0) -- (6.5,-2);
\end{tikzpicture}
\\
\begin{itemize}
\item Magnetic field out of diagram
\item Atoms in the sample are ionised
\item Directed in a narrow beam into the magnetic field
\item Radius of path depends on specific charge
\end{itemize}
$$r=\dfrac{mv}{BQ}$$
$$r\propto\dfrac{1}{\frac{q}{m}}$$
\subsubsection{Velocity selector}
\includegraphics[width=10cm]{velocity_selector.jpg}\\
Negative ions fall to the bottom surface, causing the top surface to have a positive charge, creating a potential difference. For an ion to continue on a straight path the force from the magnetic field up must be equal to the force from the electric field down.\\
$$F=BQv=EQ$$
$$BQv=\dfrac{V}{d}Q$$
$$v=\dfrac{V}{Bd}$$
This means that only particles of speed $v$ will travel through the gap

### 2018-02-19 Physics Notes/Further Mechanics and Thermal Physics/Circular Motion/Circular Motion.tex
Notes Update

\begin{document}
\begin{center}
\underline{\huge Circular Motion}
\end{center}
\section{Introduction to circular motion}
For an object to be going in a circle a force is needed to accelerate the object. This is because in a circle an object is constantly changing velocity, even if it isn't changing speed.\\
Centripetal force is not an extra force, it is just the name for a resultant force.\\
For an object to be deflected towards the centre, there needs to be a force acting towards the centre.
\subsection{Definitions}
\begin{tabular}{|c|c|c|c|}
\hline
Quantity&Symbol&Unit&Definition\\
\hline
Period&T&Seconds,s&The time taken for an object to complete one revolution\\
\hline
Frequency&f&Hertz,hz&The number of revolutions per second\\
\hline
Angular velocity&$\omega$&$\textrm{rads}^{-1}$&The rate of change in angle(the angle covered per second)\\
\hline
\end{tabular}
\subsection{Formula not on formula book}
$\omega=\dfrac{\Delta\theta}{\Delta t}$\\
$a=\dfrac{\Delta V}{\Delta t}$\\
Don't forget application of  $F=ma$ with the equations for force
\section{Circular motion for going over a hump}
\begin{tikzpicture}[
    force/.style={>=latex,draw=black,fill=black},
    m/.style={rectangle,draw=black,fill=white,minimum size=0.3cm,thin},
]
    \node[m] (m) {};
    {[force,->]
        \draw (m.north) -- ++(0,1) node[above] {$r$};
        \draw (m.south) -- ++(0,-1) node[below] {$mg$};
    }
\end{tikzpicture}
\\
For an object going over a hump the force of the weight is in the opposite direction to the reaction force.\\
The object will leave the ground if $mg>r$.
\subsection{Calculations}
$F_c=mg-R=\dfrac{mv^2}{r}$\\
\textbf{Calculating the speed at which the car will leave the ground (R=0)}\\
$mg=\dfrac{mv_0^2}{r} \qquad V_0=\sqrt{gr}$\\
Once the car has left the ground, use SUVAT equations rather than circular motion.
\section{Roundabouts}
On a roundabout the centripetal force is provided by friction.\\
$F_r=\dfrac{mv_0^2}{r}$\\
To avoid slipping, the centripetal force must be less than the maximum possible for the car and road surface.\\
\textbf{Example question}
\textit{The maximum possible force of friction between a car and the road is 800N, given a mass of 1000kg and a radius of 5m, what is the maximum speed the car can go without slipping}
\newpage
\section{Banked motion}
N is the reaction force\\
$n\sin\theta=\dfrac{mv^2}{r}$ - Force towards the centre of notion is equal to the centripetal force\\
$N\cos\theta=mg$ - The force vertically downwards is equal to the weight\\
$\tan\theta=\dfrac{v^2}{rg}$ - Combining the equations\\
$v^2=gr\tan\theta$ - Re-write
\section{Forces in a circle}
\begin{tikzpicture}[
    force/.style={>=latex,draw=black,fill=black},
    m/.style={rectangle,draw=black,fill=lightgray,minimum size=0.3cm,thin},
]
    \node[m] (m) at (0,-2) {};
    \node[m] (m2) at (0,2) {};
    \node[m] (m3) at (2,0) {};
    {[force,->]
        \draw (m.north) -- ++(0,0.7) node[above] {$R$};
        \draw (m.south) -- ++(0,-0.7) node[right] {$mg$};
    }
    {[force,->]
        \draw (m2.south) -- ++(0,-0.7) node[below] {$R+mg$};
    }
    {[force,->]
        \draw (m3.west) -- ++(-0.7,0) node[left] {$R$};
        \draw (m3.south) -- ++(0,-0.7) node[below right] {$mg$};
    }

>>>>

\begin{document}
\begin{center}
\underline{\huge Circular Motion}
\end{center}
\section{Introduction to circular motion}
For an object to be going in a circle a force is needed to accelerate the object. This is because in a circle an object is constantly changing velocity, even if it isn't changing speed.\\
Centripetal force is not an extra force, it is just the name for a resultant force.\\
For an object to be deflected towards the centre, there needs to be a force acting towards the centre.
\subsection{Definitions}
\begin{tabular}{|c|c|c|c|}
\hline
Quantity&Symbol&Unit&Definition\\
\hline
Period&T&Seconds,s&The time taken for an object to complete one revolution\\
\hline
Frequency&f&Hertz,hz&The number of revolutions per second\\
\hline
Angular velocity&$\omega$&$\textrm{rads}^{-1}$&The rate of change in angle(the angle covered per second)\\
\hline
\end{tabular}
\subsection{Formula not on formula book}
$$\omega=\frac{\Delta\theta}{\Delta t}$$
$$a=\frac{\Delta V}{\Delta t}$$
Don't forget application of  $F=ma$ with the equations for force
\section{Circular motion for going over a hump}
\begin{tikzpicture}[
    force/.style={>=latex,draw=black,fill=black},
    m/.style={rectangle,draw=black,fill=white,minimum size=0.3cm,thin},
]
    \node[m] (m) {};
    {[force,->]
        \draw (m.north) -- ++(0,1) node[above] {$r$};
        \draw (m.south) -- ++(0,-1) node[below] {$mg$};
    }
\end{tikzpicture}
\\
For an object going over a hump the force of the weight is in the opposite direction to the reaction force.\\
The object will leave the ground if $mg>r$.
\subsection{Calculations}
$$F_c=mg-R=\frac{mv^2}{r}$$
\textbf{Calculating the speed at which the car will leave the ground (R=0)}\\
$$mg=\frac{mv_0^2}{r} \qquad V_0=\sqrt{gr}$$
Once the car has left the ground, use SUVAT equations rather than circular motion.
\section{Roundabouts}
On a roundabout the centripetal force is provided by friction.\\
$$F_r=\frac{mv_0^2}{r}$$
To avoid slipping, the centripetal force must be less than the maximum possible for the car and road surface.\\
\textbf{Example question}
\textit{The maximum possible force of friction between a car and the road is 800N, given a mass of 1000kg and a radius of 5m, what is the maximum speed the car can go without slipping}
\newpage
\section{Banked motion}
N is the reaction force\\
$n\sin\theta=\dfrac{mv^2}{r}$ - Force towards the centre of notion is equal to the centripetal force\\
$N\cos\theta=mg$ - The force vertically downwards is equal to the weight\\
$\tan\theta=\dfrac{v^2}{rg}$ - Combining the equations\\
$v^2=gr\tan\theta$ - Re-write
\section{Forces in a circle}
\begin{tikzpicture}[
    force/.style={>=latex,draw=black,fill=black},
    m/.style={rectangle,draw=black,fill=lightgray,minimum size=0.3cm,thin},
]
    \node[m] (m) at (0,-2) {};
    \node[m] (m2) at (0,2) {};
    \node[m] (m3) at (2,0) {};
    {[force,->]
        \draw (m.north) -- ++(0,0.7) node[above] {$R$};
        \draw (m.south) -- ++(0,-0.7) node[right] {$mg$};
    }
    {[force,->]
        \draw (m2.south) -- ++(0,-0.7) node[below] {$R+mg$};
    }
    {[force,->]
        \draw (m3.west) -- ++(-0.7,0) node[left] {$R$};
        \draw (m3.south) -- ++(0,-0.7) node[below right] {$mg$};
    }
